### 2021-10-27 05_report/Manuscript_results.Rmd
PLGG Economic model
Initial commit

---
title: "Manuscript results"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r required packages}
require(tidyverse)    # for data manipulation
require(formattable)  # for formatting tables
require(knitr)        # for saving tables
require(scales)       # for ggplot axis labels
library(gridExtra)    # for gridded plots 
require(survival)     # for survival analysis
require(survminer)    # for survival plotting

```

## Data location

```{r}

fileloc <- here::here("04_analysis","output")
#fileloc <- "/Users/davidrios/OneDrive - SickKids/tomacbook"

```

The files are spread out over 5 folders.

## Table 2 & Abstract

Coding the following:

Table 2:Estimates of life-years after diagnosis, QALYs (Discounted) and Costs (Discounted) Estimates reported with 95% Confidence Interval in brackets. All costs are given in Canadian dollars.

Abstract results

```{R}
# model results 
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)
# turning into dataframe
df_model  <- map_df(res_model, readRDS)

# creating a key in the same format as the one in correct_surv
df_model <- df_model %>%
  mutate(key = paste0(sim,"_",rr))

# fixing survival for everyone who doesn't have a correct survival time
df_model <- df_model %>% 
    mutate( Value = ifelse(Variable == "LY", 80-Value, Value))

# no longer need key 
df_model <- df_model %>%
  select(-key)

df_delta <- df_model %>% 
  spread(intervention, Value) %>% 
  mutate(Delta = Norad_Fused-Rad_Fused) %>% 
  select(-Norad_Fused,-Rad_Fused)

df_model <- df_model %>% 
  group_by(Variable, Type, subset, intervention, discount, rr) %>% 
  summarise( mean = mean(Value),
             q025 = quantile(Value, 0.025),
             q975 = quantile(Value, 0.975)) %>% ungroup()
df_model <-df_model %>% 
  mutate(mean = case_when(Variable == "Cost"~ paste0(currency(mean,
                                                              symbol = "\\$" ,digits = 0)),
                          Variable != "Cost"~ paste0(round(mean, 2))),
         q025 = case_when(Variable == "Cost"~ paste0(currency(q025, symbol = "\\$" ,digits = 0)),
                          Variable != "Cost"~ paste0(round(q025, 2))),
        q975 = case_when(Variable == "Cost"~ paste0(currency(q975, symbol = "\\$" ,digits = 0)),
                          Variable != "Cost"~ paste0(round(q975, 2))                  
                          )
         ) %>% 
  mutate( stat = paste0(mean, " (", q025, ";",q975,")"))

df_model1 <- df_model %>% 
  select(-mean, - q025, -q975)

df_delta1 <- df_delta %>% rename(Value = Delta) %>% 
  group_by(Variable, Type, subset, discount, rr) %>% 
  summarise( mean = mean(Value),
             q025 = quantile(Value, 0.025),
             q975 = quantile(Value, 0.975)) %>% ungroup() %>% 
  mutate(mean = case_when(Variable == "Cost"~ paste0(currency(mean, symbol = "\\$" ,digits = 0)),
                          Variable != "Cost"~ paste0(round(mean, 2))),
         q025 = case_when(Variable == "Cost"~ paste0(currency(q025, symbol = "\\$" ,digits = 0)),
                          Variable != "Cost"~ paste0(round(q025, 2))),
        q975 = case_when(Variable == "Cost"~ paste0(currency(q975, symbol = "\\$" ,digits = 0)),
                          Variable != "Cost"~ paste0(round(q975, 2))

                          )
         ) %>% 
  mutate( stat = paste0(mean, " (", q025, ";",q975,")")) %>% 
    select(-mean, - q025, -q975)

df_deltacea <- df_delta %>% 
    filter(Type == "Total" | Type == "QALY", subset == "all", discount == 1 ) %>% select(-Type)  %>% spread(Variable, Delta)

df_model2 <- df_model1 %>% 
  bind_rows(mutate( df_delta1, intervention = "delta")) %>% 
    mutate(Variable = factor(Variable, levels = c("LY","QALY","Cost")),
         Type = factor(Type, levels = c("LY","QALY","Total","PLGG","AE"))
         ) %>% 
  arrange(Variable, Type, discount, rr) %>% 
  spread(intervention, stat)

df_model3 <- df_model2 %>% 
  arrange(subset, rr, Variable, Type, discount )

tb2 <-df_model3  %>%
  filter(subset == "all") %>% 
  filter( !(Variable == "Cost" & discount == 0),
          !(Variable == "QALY" & discount == 0) ) %>%
  select(-Variable, - subset,-discount  ) %>% 
  rename(Variable = Type,
         Delta = delta,
         intervention = Norad_Fused,
         control = Rad_Fused)

```

**Table 2**:

```{r}
library(pander)
library(forcats)

# kable(tb2)
tb2a <- tb2 %>% 
  mutate(Variable = fct_recode(Variable, `Life-years` = "LY", 
                               QALY = "QALY",
                               `Total Cost` = "Total")) %>%
  select(Variable, intervention, control, Delta , rr)

pander(tb2a)
```

## Supplementary Table 1

Supplementary Table 1: Estimates of life-years after diagnosis, QALYs (Discounted) and Costs (Discounted), for fused patients, and fused patients who progressed.

```{r}

sm_tb1 <-df_model3  %>%
  filter(subset != "all") %>% 
  filter( !(Variable == "Cost" & discount == 0),
          !(Variable == "QALY" & discount == 0) ) %>%
  select(-Variable,-discount  ) %>% 
  rename(Variable = Type,
         Delta = delta,
         intervention = Norad_Fused,
         control = Rad_Fused)

sm_tb1 %>%
  kable()
```

## Figure 4 Cost-effectiveness plane

Figure 4: Cost-effectiveness plane indicating the incremental cost and incremental effectiveness estimates related to molecular testing for each probabilistic simulation assuming both a no-radiation (n = 1,000) and radiation benefit (n = 1,000)

```{r}
df_deltacea <- df_delta %>% 
    filter(Type == "Total" | Type == "QALY", subset == "all", discount == 1 ) %>% select(-Type)  %>% spread(Variable, Delta)

df_deltacea1 <- df_deltacea %>% 
  mutate(rr =factor(rr, levels = c(0,1), labels = c("No radiation benefit","Radiation benefit")),
         type = "all")
df_means <- df_deltacea %>% 
  group_by(rr) %>% 
  summarise(Cost = mean(Cost),
            QALY = mean(QALY)) %>% 
  mutate(type = "Mean Values") %>% 
  mutate(rr1 = factor(rr, levels = c(0,1), labels = c("No radiation benefit","Radiation benefit") ))
toplot_plane <-df_deltacea1 %>% 
  ggplot() + 
  geom_point(aes(x = QALY, y = Cost), alpha = 0.75 , size = 0.6) + 
  # geom_point(aes(x = QALY, y = Cost), color = "black", data = df_means,shape = 2,show.legend = TRUE, size = 3) + 
  geom_hline(yintercept = 0) + 
  geom_vline(xintercept = 0) + 
  theme_bw() + 
  theme(legend.title = element_blank()) +
  ylab("Discounted Total Costs") +  
  xlab("Discounted QALYs") + 
  labs(color = "", shape = "Mean") +
  scale_y_continuous(labels = dollar_format(prefix="$")) +
  facet_wrap(~ rr) + 
  theme( panel.grid.major = element_blank(),
panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))

toplot_plane

```

# Figure 2

Figure 2: Median cumulative incidence of all-cause mortality for all patients assuming n o radiation benefit or with radiation benefit.

```{r}
res_os <- list.files(paste0(fileloc, "/res_OS"),full.names = TRUE)
df_OS <- map_df(res_os, readRDS)

df_med <- df_OS %>%
  group_by(cycle, type, intervention, rr) %>%
  summarise(median = median(OS)) %>% ungroup() %>%
   mutate(rr = ifelse(rr == 0, "No radiation benefit", "Radiation benefit")) %>%
    mutate(intervention  = ifelse(intervention == "Norad_Fused", "Intervention","Control"))

df_delta <- df_OS %>%
  spread(intervention, OS) %>%
  mutate(delta = (1-Norad_Fused) - (1-Rad_Fused)) %>%
  select(-Norad_Fused, - Rad_Fused) %>%
  group_by(cycle, type, rr) %>%
  summarise(median = median(delta),
            q025 = quantile(delta, 0.025),
            q975 = quantile(delta, 0.975)) %>% ungroup() %>% 
  mutate(rr = ifelse(rr == 0, "No radiation benefit", "Radiation benefit"))

rm(df_OS)

plot1 <- filter(df_med,type == "OS") %>%
  ggplot() + 
  geom_line(aes(x = cycle, y = 1-median, linetype = intervention)) +
  facet_wrap(~rr) +
  theme_bw() + 
  ylab("\n Cumulative Incidence, all-cause mortality") + 
  xlab("Time from diagnosis, Years")  + 
    theme(legend.title =element_blank(),legend.position="top") +
  theme( panel.grid.major = element_blank(),
panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))

plot2_delta <- filter(df_delta,type == "OS" ) %>%
  mutate(t1 = "Median (95% CI)") %>% 
  ggplot() + 
  geom_line(aes(x = cycle, y = median, color = t1)) + 
  geom_ribbon(aes(x = cycle, ymin = q025, ymax = q975, fill = t1), alpha = 0.3,show.legend = T) +
  facet_wrap(~rr) + theme_bw() +
  ylab("Delta (Intervention - Control) \n Cumulative Incidence, all-cause mortality") + 
    xlab("Time from diagnosis, Years") + 
  theme(legend.title =element_blank(),legend.position="top")  +
  scale_color_manual(values = "black") +
  scale_fill_manual( values = "black") +
  theme( panel.grid.major = element_blank(),
panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))

fig2 <-grid.arrange(plot1, plot2_delta)

fig2
```

In text results

```{r}
df_med %>% 
  filter(cycle == 20) %>% mutate(median = median*100) %>% 
  spread(intervention, median) %>% 
  arrange(rr, type) %>% filter(type == "OS" | type == "Fused") %>%  
  kable(digits = 1 ) 
```

# Supp Fig 1

Supplementary Figure 1: Cumulative incidence of all-cause mortality for all fused patients, and for fused patients who progressed

```{r}
plot2SM <-filter(df_delta,type != "OS" ) %>%
  mutate(type = factor(type, levels = c("Fused","FusedProg"),
                       labels = c("Fused patients", "Fused patients that progressed"))) %>% 
 mutate(t1 = "Median (95% CI)") %>% 
  ggplot() + 
  geom_line(aes(x = cycle, y = median, color = t1)) + 
  geom_ribbon(aes(x = cycle, ymin = q025, ymax = q975, fill = t1), alpha = 0.3,show.legend = T) +
  facet_wrap(~rr + type,) + theme_bw() +
  ylab("Delta (Intervention - Control) \n Cumulative Incidence, all-cause mortality") + 
    xlab("Time from diagnosis, Years") + 
  theme(legend.title =element_blank(),legend.position="top")  +
  scale_color_manual(values = "black") +
  scale_fill_manual( values = "black")  +
  theme( panel.grid.major = element_blank(),
panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))

plot2SM

rm(fig2)
```

## Figure 3

Create figure 3

First accounting for mortality

```{r eval=T, message=FALSE, warning=FALSE, include=FALSE}
res_CI <- list.files(paste0(fileloc, "/res_CI"),full.names = TRUE)
res_OS <- list.files(paste0(fileloc, "/res_OS"),full.names = TRUE)

res_CI_s <- list.files(paste0(fileloc, "/res_CI"),full.names = FALSE)
res_OS_s <- list.files(paste0(fileloc, "/res_OS"),full.names = FALSE)

res_OS <- res_OS[order(res_OS_s)]
res_CI <- res_CI[order(res_CI_s)]

res_short <- list.files(paste0(fileloc, "/res_CI"),full.names = F)
res_short <- res_short[order(res_OS_s)]
v_n <-c( "neurologic", "auditory", "visual", "stroke", "cardiovascular", "SN")

## first read in first thing

for( i in seq_along(res_short)){
  if(i %% 10 == 0){message(i/length(res_short))}
tempCI <- readRDS(res_CI[i])
tempOS <- readRDS(res_OS[i]) %>% rename(time = cycle) %>% filter(type == "OS")
tempOS <-tempOS %>% group_by(type, sim , intervention, rr) %>% 
  mutate(OS_t1 = lag(OS, 1,1)) %>% 
  select(-type)

tempCI <- tempCI %>% 
  left_join(tempOS, by = c("time", "sim", "intervention", "rr"))

calcCI <- tempCI %>% 
  filter(state %in% v_n) %>%
    mutate(ci = (ceiling(ci*100000)*ceiling(OS*100000))/100000) %>% 
    mutate(ci = ceiling(ci*100000),
           OS = ceiling(OS*100000),
           OS_t1= ceiling(OS_t1*100000)/100000) %>%
  group_by(state, intervention, rr ) %>%
  mutate(ci = c(0,diff(ci,1))) %>% 
  mutate(mcf =  ci/OS,
         CI = mcf*OS_t1
         ) %>% 
  mutate(cumCI = cumsum(CI))  %>%
  select(state, intervention ,rr, time,cumCI, sim ) %>% ungroup()

for(state1 in v_n){
  dir.create(path = here::here("05_report","1newCIcalc",state1),recursive = T)

     filter(calcCI, state == state1) %>%
      saveRDS(file = here::here("05_report","1newCIcalc",state1,"/",paste0(state1,i,"BIG.RDS")))

}

}

```

```{r}
store_toplot <- vector(mode = "list", length = length(v_n))

for(j in seq_along(v_n)){
  message(j)

  list_files_instate <- list.files(here::here("05_report","1newCIcalc",v_n[j]),full.names = TRUE)

big1   <- map_df(list_files_instate, readRDS)

store_toplot[[j]] <- big1 %>% 
  group_by(intervention, state,rr) %>%
  group_by(intervention, state,rr) %>%
  mutate(cumCI = ifelse(is.nan(cumCI) , NA, cumCI  )) %>%
  fill(cumCI,.direction = "down") %>%
  group_by(intervention, state, rr,  time  ) %>%
summarise(median = median(cumCI),
          q025 = quantile(cumCI, 0.025),
          q975 = quantile(cumCI, 0.975)) %>% ungroup()

}

unlink(here::here("05_report","1newCIcalc"),recursive = T)

```

```{r}

toplotCI <- do.call(rbind,store_toplot)

plotCI <- toplotCI %>% ungroup() %>%
    mutate(intervention  = ifelse(intervention == "Norad_Fused", "Intervention","Control")) %>% 
    mutate(rr= ifelse(rr == 1, "Cherlow, et al.","No radiation benefit")) %>%
  ggplot() + 
  geom_line(aes(x = time, y = median, color = intervention)) +   geom_ribbon(aes(x = time, ymin = q025, ymax = q975 , fill = intervention), alpha = 0.3,show.legend = T) +
 facet_grid(rows =  vars(state) , cols =  vars(rr) ) +
theme_bw() +
  ylab("Cumulative incidence, Treatment related AE") + 
    xlab("Time from diagnosis, Years") + 
  theme( panel.grid.major = element_blank(),
panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))

plotCI
```

```{r}

dir.create(here::here("05_report","tempfold1"))
res_CI <- list.files(paste0(fileloc, "/res_CI"),full.names = TRUE)
res_OS <- list.files(paste0(fileloc, "/res_OS"),full.names = TRUE)

res_CI_s <- list.files(paste0(fileloc, "/res_CI"),full.names = FALSE)
res_OS_s <- list.files(paste0(fileloc, "/res_OS"),full.names = FALSE)

res_OS <- res_OS[order(res_OS_s)]
res_CI <- res_CI[order(res_CI_s)]

res_short <- list.files(paste0(fileloc, "/res_CI"),full.names = F)
res_short <- res_short[order(res_OS_s)]

list1 <- c("stroke_gen","cardiovascular_gen","SN_gen")
list2 <- c("stroke","cardiovascular","SN")
v_n <- c(list1,list2)

## first read in first thing

for( i in 1:length(res_short)){
  if(i %% 10 == 0){message(i/length(res_short))} 
  tempCI <- readRDS(res_CI[i]) 
  tempOS <- readRDS(res_OS[i]) %>% 
    rename(time = cycle) %>% filter(type == "OS")

  tempOS <-tempOS %>% group_by(type, sim , intervention, rr) %>%
  mutate(OS_t1 = lag(OS, 1,1)) %>%
  select(-type)

tempCI <- tempCI %>%
  left_join(tempOS, by = c("time", "sim", "intervention", "rr"))

## Have TempCI joined with tempOS
## issue is that i calcualted CI = cumsum(mean(incidence*OS)) ##  mean(abs(mDur1[,ci_states[j]] - time_v[t1]) < tol)/per_alive[t1] ## cumsum(k1) ## Reversing that

calCI <-tempCI %>%
  filter(state %in% v_n) %>%
  group_by(state, sim, intervention ,rr, type) %>%arrange(time) %>%
  mutate(i = c(0, diff(ci)),
         i = (i*100000)*OS,
         ) %>% ungroup() %>%
  mutate(state = case_when(
state == list1[1] |state ==  list2[1] ~ list2[1], state == list1[2] |state ==  list2[2] ~ list2[2], state == list1[3] |state ==  list2[3] ~ list2[3]
)) %>% 
   group_by(state, sim, intervention ,rr, type, time) %>%
  mutate( i = sum(i)) %>%  ungroup() %>% 
   group_by(state, sim, intervention ,rr, type) %>%
   mutate(mcf =  i /(OS*100000),
         CI = mcf*OS_t1
         )  %>% 
    mutate(cumCI = cumsum(CI))  %>%
  select(state, intervention ,rr, time,cumCI, sim, type )

# trying to fix that cumulative incidence doesnt take into account mortality

      saveRDS(file = here::here("05_report","tempfold1",paste0(i,"BIG.RDS")),object = calCI) 
      } 
```

```{r}

file1 <- list.files(here::here("05_report","tempfold1"),full.names = T)

bigdf<-map_df(file1, readRDS)

bigdf %>% arrange(desc(cumCI))

bigdf %>% 
   mutate(rr = ifelse(rr == 0, "No radiation benefit", "Radiation benefit")) %>%
    mutate(intervention  = ifelse(intervention == "Norad_Fused", "Intervention","Control")) %>%
  group_by(state, intervention, rr, time,type) %>% 
  summarise(mean = mean(cumCI),
            q025 = quantile(cumCI,0.025),
            q975 = quantile(cumCI,0.975)) %>%

  ggplot() +
  facet_wrap(~state+ rr) + geom_line(aes(x = time, y= mean, color = intervention)) +
  geom_ribbon(aes(x = time, ymin = q025, ymax = q975, fill = intervention), alpha = 0.1)+
  theme_bw()

unlink(here::here("05_report","tempfold1"),recursive = T)

```