### 2023-12-05 04_analysis/control scenarios/Bryans SickKids data/chart_audit_survival_new_controls_firstline.Rmd
updated to targeted first line

---
title: "Survival Analysis Chart Audit Data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(lubridate)
library(survival)
library(flexsurv)
library(survminer)
library(here)
library(darthtools)
library(data.table)
```

- date last FU CCO gone
- age_diag same as age_diag_years

```{r}
loc_here <- getwd()
# df0 <- read_csv(paste0(loc_here, "02_data/PLGG_targeted_chartaudit_form.csv"), skip=1)
# df <- read_csv("plgg_cea_controls.csv")
df <- read_csv("PLGG_targetedtherapy_CEA_Petros_Nov2023_firstline.csv")
df$status <- ifelse(df$secondprogress   == 1 | df$death == 1, 1 ,0)
df$time <- df$TimetoOutcome_days /365 # change time in days to in years
```

Kaplan-Meier estimates of PFS

```{r}
# Kaplan-Meier fit for PFS
KM_PFS <- survfit(Surv(time = time, event = status) ~ 1, data = df)
# plot KM for PFS
ggsurvplot(
  KM_PFS,
  data       = df,
  size       = 1,                # change line size
  conf.int   = FALSE,            # Add confidence interval
  pval       = F,                # Add p-value
  risk.table = F,                # Add risk table
  risk.table.height = 0.25,      # Useful to change when you have multiple groups
  ggtheme    = theme_bw(),       # Change ggplot2 theme
  xlab       = 'Time in years', # Change X-axis label
  title      = "Survival curve for PFS",
  subtitle   = "Based on Kaplan-Meier estimates"
)
```

# Fit survival curves:

*Control arm:*

```{r}
times <- 0:80
mods <- c("exp", "weibull", "gamma", "lnorm", "llogis", "gengamma", "rps")
fit_control <- fit.fun(time="time", status="status", mods=mods, 
                       data=df %>% as.data.frame(), times=times,
                       extrapolate = T) 
a <- fit_control$plot
a$plot <- a$plot + labs(subtitle="Control arm (n=34)") + scale_x_continuous(breaks=seq(0,80,10))
a$plot
```

Choose best-fitting survival model:

```{r}
GoF_control <- data.frame(AIC = fit_control$AIC, BIC = fit_control$BIC)
choose_PFS_AIC_control <- rownames(GoF_control)[which.min(GoF_control$AIC)]; choose_PFS_AIC_control
choose_PFS_BIC_control <- rownames(GoF_control)[which.min(GoF_control$BIC)]; choose_PFS_BIC_control
```

```{r}
mods <- c("gamma", "exp")
fit_control <- fit.fun(time="time", status="status", mods=mods, 
                       data=df %>% as.data.frame(), times=times,
                       extrapolate = T) 
a1 <- fit_control$plot
a1$plot <- a1$plot + labs(subtitle="Control arm (n=34), best-fitting curve") + 
  scale_x_continuous(breaks=seq(0,80,10))
a1$plot
```

Calculate transition probabilities (prog 1 to prog 2):

```{r}
best_model_control <- fit_control$model.objects$models$`Exponential`
surv_probs_control <- surv_prob(best_model_control, times = times) # surv probs (not having 2nd prog) from prog 1
trans_probs_control <- trans_prob(surv_probs_control) # transition probs from prog 1 to prog 2
```

Bootstrap for microsim:

```{r}
N_sim <- 100
fitted_PFS_chemo <- list()
set.seed(1)
for (i in 1:N_sim) {
  ids <- sample(1:nrow(df), replace=T)
  PFS_data_boot <- df[ids,]
  fit_PFS <- fit.fun(time  = "time", status = "status", 
                     data = PFS_data_boot %>% as.data.frame(),
                     times = times, extrapolate = T, mods=c("exp"))
  fitted_PFS_chemo[[i]] <- fit_PFS$model.objects$models$Exponential
  print(i)
}
save(fitted_PFS_chemo, file="fitted_PFS_control_sk_firstline.RData")
```

```{r}

```