### 2023-12-05 04_analysis/control scenarios/Erics firstline RCT/first_line_paper_survival.Rmd

---
title: "PFS analysis of paper - first line"
output:
  pdf_document: default
  html_document: default
---

Paper:

https://pubmed.ncbi.nlm.nih.gov/37733309/#:~:text=Conclusions%3A%20Among%20pediatric%20patients%20with,chemotherapy%20as%20first%2Dline%20therapy

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, eval = TRUE)
rm(list = ls())     # clear memory (removes all the variables from the workspace)
options(scipen=999) # disable scientific notation
```

```{r, warning=F, message=F}
if (!require('pacman')) install.packages('pacman'); library(pacman) # use this package to conveniently install other packages
# load (install if required) packages from CRAN
p_load("here", "dplyr", "devtools", "gems", "flexsurv", "flexsurvcure", "survHE", "survminer", "ggplot2", "msm", "igraph", "mstate", "reshape2", "knitr", "abind", "dampack", "data.table", "tm")

# load (install if required) packages from GitHub
# install_github("DARTH-git/darthtools", force = TRUE) Uncomment if there is a newer version
p_load_gh("DARTH-git/darthtools")
```

# 04 Digitized Data

Use the function `digitise()` to translate the digitised OS and PFS data into patient level information.

## Control arm

```{r}
time_interval <- seq(0, 32, by = 2)
at_risk <- c(37,25,19,17,10,7,4,2,1,1,1,1,0,0,0,0,0)
surv_data <- read.csv("out_put_alanNejmoa.png_Auto_Digitized.csv", header = T)
surv_data <- surv_data[surv_data$curve==2,]
surv_data$St <- surv_data$St/100
surv_data <- data.frame(k = 1:nrow(surv_data),
                        Time = surv_data$time,
                        Sk = surv_data$St)
write.table(surv_data, file = "first_line_PFS_curve_control.txt", row.names = F)
surv_data <- read.table("first_line_PFS_curve_control.txt", header = T)
surv_times <- surv_data$Time

n_risk_table <- create_at_risk_table(survival_time = surv_times,
                                     Years = time_interval,
                                     AtRisk = at_risk)
write.table(n_risk_table, "first_line_PFS_risktable_control.txt", row.names = F)
```

```{r, results="hide"}
# Create IPD and KM data for the PFS curves (there have been 99 events in the cohort)
digitise(surv_inp   = "first_line_PFS_curve_control.txt", 
         nrisk_inp  = "first_line_PFS_risktable_control.txt", 
         km_output  = "first_line_PFS_KM_control.txt", 
         ipd_output = "first_line_PFS_IPD_control.txt")

# Link the IPD files across the two arms of the trial for OS and PFS
IPD_PFS <- make.ipd(ipd_files = c("first_line_PFS_IPD_control.txt"), ctr = 1,
                    var.labs  = c("time","event","arm"))
IPD_PFS <- IPD_PFS %>% select(-arm)
```

```{r}
library(survival)
library(survminer)
# Kaplan-Meier fit for PFS
KM_PFS <- survfit(Surv(time = time, event = event) ~ 1, data = IPD_PFS)
# plot KM for PFS
ggsurvplot(
  KM_PFS,
  data       = IPD_PFS,
  size       = 1,                # change line size
  conf.int   = FALSE,            # Add confidence interval
  pval       = F,                # Add p-value
  risk.table = T,                # Add risk table
  risk.table.height = 0.25,      # Useful to change when you have multiple groups
  break.x.by = 2,
  ggtheme    = theme_bw(),       # Change ggplot2 theme
  xlab       = 'Time in months', # Change X-axis label
  title      = "Survival curve for PFS, Control arm",
  subtitle   = "Based on Kaplan-Meier estimates"
)
```

```{r}
surv_data$Time <- surv_data$Time/12 # change time in months to years
# times <- seq(0,5,by=1/12)
times <- 0:80
mods <- c("exp", "weibull", "gamma", "lnorm", "llogis", "gompertz" ,"rps")
fit_PFS  <- fit.fun(time = "time", status = "event", data = IPD_PFS, times = times,
                    extrapolate = T, mods = mods) 
```

```{r}
GoF_PFS <- data.frame(AIC = fit_PFS$AIC, BIC = fit_PFS$BIC)
choose_PFS <- rownames(GoF_PFS)[which.min(GoF_PFS$AIC)]; choose_PFS
choose_PFS <- rownames(GoF_PFS)[which.min(GoF_PFS$BIC)]; choose_PFS
```

```{r}
mods=c("lnorm", "exp")
fit_PFS  <- fit.fun(time = "time", status = "event", data = IPD_PFS, times = times,
                    extrapolate = T, mods=mods) 
```

```{r}
N_sim <- 100
fitted_PFS_chemo <- list()
set.seed(1)
for (i in 1:N_sim) {
  ids <- sample(1:nrow(IPD_PFS), replace=T)
  PFS_data_boot <- IPD_PFS[ids,]
  fit_PFS <- fit.fun(time  = "time", status = "event", data = PFS_data_boot,
                     times = times, extrapolate = T, mods=c("lnorm"))
  fitted_PFS_chemo[[i]] <- fit_PFS$model.objects$models$`log-Normal`
  print(i)
}
save(fitted_PFS_chemo, file="fitted_PFS_control_firstline.RData")
```

## Targeted arm

```{r}
time_interval <- seq(0, 32, by = 2)
at_risk <- c(73,67,64,62,49,40,33,25,23,13,11,7,4,3,2,2,0)
surv_data <- read.csv("out_put_alanNejmoa.png_Auto_Digitized.csv", header = T)
surv_data$St <- surv_data$St/100
surv_data <- surv_data[surv_data$curve==1,]
surv_data <- data.frame(k = 1:nrow(surv_data),
                        Time = surv_data$time,
                        Sk = surv_data$St)
write.table(surv_data, file = "first_line_PFS_curve_targeted.txt", row.names = F)
surv_data <- read.table("first_line_PFS_curve_targeted.txt", header = T)
surv_times <- surv_data$Time

n_risk_table <- create_at_risk_table(survival_time = surv_times,
                                     Years = time_interval,
                                     AtRisk = at_risk)
write.table(n_risk_table, "first_line_PFS_risktable_targeted.txt", row.names = F)
```

```{r, results="hide"}
# Create IPD and KM data for the PFS curves (there have been 99 events in the cohort)
digitise(surv_inp   = "first_line_PFS_curve_targeted.txt",
         nrisk_inp  = "first_line_PFS_risktable_targeted.txt",
         km_output  = "first_line_PFS_KM_targeted.txt",
         ipd_output = "first_line_PFS_IPD_targeted.txt")

# Link the IPD files across the two arms of the trial for OS and PFS
IPD_PFS <- make.ipd(ipd_files = c("first_line_PFS_IPD_targeted.txt"), ctr = 1,
                    var.labs  = c("time","event","arm"))
IPD_PFS <- IPD_PFS %>% select(-arm)
```

```{r}
library(survival)
library(survminer)
# Kaplan-Meier fit for PFS
KM_PFS <- survfit(Surv(time = time, event = event) ~ 1, data = IPD_PFS)
# plot KM for PFS
ggsurvplot(
  KM_PFS,
  data       = IPD_PFS,
  size       = 1,                # change line size
  conf.int   = FALSE,            # Add confidence interval
  pval       = F,                # Add p-value
  risk.table = T,                # Add risk table
  risk.table.height = 0.25,      # Useful to change when you have multiple groups
  break.x.by = 2,
  ggtheme    = theme_bw(),       # Change ggplot2 theme
  xlab       = 'Time in months', # Change X-axis label
  title      = "Survival curve for PFS, Targeted arm",
  subtitle   = "Based on Kaplan-Meier estimates"
)
```

```{r}
surv_data$Time <- surv_data$Time/12
# times <- seq(0,5,by=1/12)
times <- 0:80
mods <- c("exp", "weibull", "gamma", "lnorm", "llogis", "gompertz" ,"rps")
fit_PFS  <- fit.fun(time = "time", status = "event", data = IPD_PFS, times = times,
                    extrapolate = T, mods = mods)
```

```{r}
GoF_PFS <- data.frame(AIC = fit_PFS$AIC, BIC = fit_PFS$BIC)
choose_PFS <- rownames(GoF_PFS)[which.min(GoF_PFS$AIC)]; choose_PFS
choose_PFS <- rownames(GoF_PFS)[which.min(GoF_PFS$BIC)]; choose_PFS
```

```{r}
mods=c("exp", "llogis")
fit_PFS  <- fit.fun(time = "time", status = "event", data = IPD_PFS, times = times,
                    extrapolate = T, mods=mods)
```

```{r}
N_sim <- 100
fitted_PFS_targeted <- list()
set.seed(1)
for (i in 1:N_sim) {
  ids <- sample(1:nrow(IPD_PFS), replace=T)
  PFS_data_boot <- IPD_PFS[ids,]
  fit_PFS <- fit.fun(time  = "time", status = "event", data = PFS_data_boot,
                     times = times, extrapolate = T, mods=c("exp"))
  fitted_PFS_targeted[[i]] <- fit_PFS$model.objects$models$Exponential
  print(i)
}
save(fitted_PFS_targeted, file="fitted_PFS_targeted_firstline.RData")
```