### 2024-08-23 05_report/Manuscript_results_temp1.Rmd
updated

---
title: "Manuscript Results"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning=F, message=F)
#rm(list = ls())
options(scipen=999)
```

```{r required packages}
require(tidyverse)    # for data manipulation
require(formattable)  # for formatting tables
require(knitr)        # for saving tables
require(scales)       # for ggplot axis labels
library(gridExtra)    # for gridded plots 
require(survival)     # for survival analysis
require(survminer)    # for survival plotting
require(kableExtra)
index <- c(1:50)
nn <- length(index)
loc_here <- paste0(here::here(),"/")
source(paste0(loc_here,"05_report/report_functions.R"))
```

# Base case

## Scenario #1a

- Combo vs. chemo as first-line treatment
- Lifetime duration of combo therapy 
- 0% radiation at progression (Post-2014 SickKids data indicates that no patients who received chemo and progressed were radiated) 
- 100% combo price
- Discount rate: 1.5 
- Data source for controls & exposed: Investigator assessment - Bouffet et al., NEJM 2023. 
- **Post-progression risk for subsequent progression = pre-progression risk for first progression in combo arm 
- 10,000 individuals
- 50 simulation runs

\newpage

### Investigator assessment, Targeted curve – Exponential, SoC curve – Exponential

```{r}
scen <- paste0("basecase/","1a")
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase exp lifetime investigator combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Investigator assessment, Targeted curve – Log-Normal, SoC curve – Log-Normal

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase lnorm lifetime investigator combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Investigator assessment, Targeted curve – Log-Normal, SoC curve – Exponential

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase lnorm exp lifetime investigator combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Independent reviewer, Targeted curve – Exponential, SoC curve – Log-Normal

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase exp lifetime combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Independent reviewer, Targeted curve – Log-Normal, SoC curve – Log-Normal

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase lnorm lifetime combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

## Scenario #1c

- Combo vs. chemo as first-line treatment
- Lifetime duration of combo therapy
- 0% radiation at progression (Post-2014 SickKids data indicates that no patients who received chemo and progressed were radiated)
- **50% combo price reduction**
- Discount rate: 1.5
- Data source for controls & exposed: Investigator assessment - Bouffet et al., NEJM 2023.
- **Post-progression risk for subsequent progression = pre-progression risk for first progression in combo arm
- 10,000 individuals
- 50 simulation runs

\newpage

### Investigator assessment, Targeted curve – Exponential, SoC curve – Exponential

```{r}
scen <- paste0("basecase/","1c")
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase exp lifetime investigator combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Investigator assessment, Targeted curve – Log-Normal, SoC curve – Log-Normal

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase lnorm lifetime investigator combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Investigator assessment, Targeted curve – Log-Normal, SoC curve – Exponential

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase lnorm exp lifetime investigator combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Independent reviewer, Targeted curve – Exponential, SoC curve – Log-Normal

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase exp lifetime combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Independent reviewer, Targeted curve – Log-Normal, SoC curve – Log-Normal

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase lnorm lifetime combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

## Scenario #1d

- Combo vs. chemo as first-line treatment
- Lifetime duration of combo therapy
- 0% radiation at progression (Post-2014 SickKids data indicates that no patients who received chemo and progressed were radiated)
- **75% combo price reduction**
- Discount rate: 1.5
- Data source for controls & exposed: Investigator assessment - Bouffet et al., NEJM 2023.
- **Post-progression risk for subsequent progression = pre-progression risk for first progression in combo arm
- 10,000 individuals
- 50 simulation runs

\newpage

### Investigator assessment, Targeted curve – Exponential, SoC curve – Exponential

```{r}
scen <- paste0("basecase/","1d")
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase exp lifetime investigator combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Investigator assessment, Targeted curve – Log-Normal, SoC curve – Log-Normal

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase lnorm lifetime investigator combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Investigator assessment, Targeted curve – Log-Normal, SoC curve – Exponential

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase lnorm exp lifetime investigator combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Independent reviewer, Targeted curve – Exponential, SoC curve – Log-Normal

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase exp lifetime combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Independent reviewer, Targeted curve – Log-Normal, SoC curve – Log-Normal

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase lnorm lifetime combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

## Scenario #1e

- Combo vs. chemo as first-line treatment
- Lifetime duration of combo therapy
- 0% radiation at progression (Post-2014 SickKids data indicates that no patients who received chemo and progressed were radiated)
- 100% combo price
- Discount rate: 1.5
- **Data source for controls: n=34 SickKids controls**
- **Data source for exposed: Investigator assessment - Bouffet et al., NEJM 2023.**
- **Post-progression risk for subsequent progression = pre-progression risk for first progression in combo arm
- 10,000 individuals
- 50 simulation runs

\newpage

### Targeted curve – Exponential, SoC curve – Exponential

```{r}
scen <- paste0("basecase/","1e")
fileloc <- here::here("04_analysis","output","scenarios",scen, "sickkids control invest exp lifetime combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Targeted curve – Log-Normal, SoC curve – Exponential

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "sickkids control invest lnorm lifetime combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

# 0% discount rate

## Scenario #1a

- Combo vs. chemo as first-line treatment
- Lifetime duration of combo therapy 
- 0% radiation at progression (Post-2014 SickKids data indicates that no patients who received chemo and progressed were radiated) 
- 100% combo price
- Discount rate: 0 
- Data source for controls & exposed: Investigator assessment - Bouffet et al., NEJM 2023. 
- **Post-progression risk for subsequent progression = pre-progression risk for first progression in combo arm 
- 10,000 individuals
- 50 simulation runs

\newpage

### Investigator assessment, Targeted curve – Exponential, SoC curve – Exponential

```{r}
scen <- paste0("0_dr/","1a")
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase exp lifetime investigator combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Investigator assessment, Targeted curve – Log-Normal, SoC curve – Log-Normal

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase lnorm lifetime investigator combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Investigator assessment, Targeted curve – Log-Normal, SoC curve – Exponential

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase lnorm exp lifetime investigator combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Independent reviewer, Targeted curve – Exponential, SoC curve – Log-Normal

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase exp lifetime combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Independent reviewer, Targeted curve – Log-Normal, SoC curve – Log-Normal

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase lnorm lifetime combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

## Scenario #1c

- Combo vs. chemo as first-line treatment
- Lifetime duration of combo therapy
- 0% radiation at progression (Post-2014 SickKids data indicates that no patients who received chemo and progressed were radiated)
- **50% combo price reduction**
- Discount rate: 0
- Data source for controls & exposed: Investigator assessment - Bouffet et al., NEJM 2023.
- **Post-progression risk for subsequent progression = pre-progression risk for first progression in combo arm
- 10,000 individuals
- 50 simulation runs

\newpage

### Investigator assessment, Targeted curve – Exponential, SoC curve – Exponential

```{r}
scen <- paste0("0_dr/","1c")
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase exp lifetime investigator combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Investigator assessment, Targeted curve – Log-Normal, SoC curve – Log-Normal

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase lnorm lifetime investigator combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Investigator assessment, Targeted curve – Log-Normal, SoC curve – Exponential

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase lnorm exp lifetime investigator combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Independent reviewer, Targeted curve – Exponential, SoC curve – Log-Normal

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase exp lifetime combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Independent reviewer, Targeted curve – Log-Normal, SoC curve – Log-Normal

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase lnorm lifetime combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

## Scenario #1d

- Combo vs. chemo as first-line treatment
- Lifetime duration of combo therapy
- 0% radiation at progression (Post-2014 SickKids data indicates that no patients who received chemo and progressed were radiated)
- **75% combo price reduction**
- Discount rate: 0
- Data source for controls & exposed: Investigator assessment - Bouffet et al., NEJM 2023.
- **Post-progression risk for subsequent progression = pre-progression risk for first progression in combo arm
- 10,000 individuals
- 50 simulation runs

\newpage

### Investigator assessment, Targeted curve – Exponential, SoC curve – Exponential

```{r}
scen <- paste0("0_dr/","1d")
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase exp lifetime investigator combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Investigator assessment, Targeted curve – Log-Normal, SoC curve – Log-Normal

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase lnorm lifetime investigator combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Investigator assessment, Targeted curve – Log-Normal, SoC curve – Exponential

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase lnorm exp lifetime investigator combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Independent reviewer, Targeted curve – Exponential, SoC curve – Log-Normal

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase exp lifetime combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Independent reviewer, Targeted curve – Log-Normal, SoC curve – Log-Normal

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase lnorm lifetime combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

## Scenario #1e

- Combo vs. chemo as first-line treatment
- Lifetime duration of combo therapy
- 0% radiation at progression (Post-2014 SickKids data indicates that no patients who received chemo and progressed were radiated)
- 100% combo price
- Discount rate: 0
- **Data source for controls: n=34 SickKids controls**
- **Data source for exposed: Investigator assessment - Bouffet et al., NEJM 2023.**
- **Post-progression risk for subsequent progression = pre-progression risk for first progression in combo arm
- 10,000 individuals
- 50 simulation runs

\newpage

### Targeted curve – Exponential, SoC curve – Exponential

```{r}
scen <- paste0("0_dr/","1e")
fileloc <- here::here("04_analysis","output","scenarios",scen, "sickkids control invest exp lifetime combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Targeted curve – Log-Normal, SoC curve – Exponential

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "sickkids control invest lnorm lifetime combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

# 3% discount rate

## Scenario #1a

- Combo vs. chemo as first-line treatment
- Lifetime duration of combo therapy 
- 0% radiation at progression (Post-2014 SickKids data indicates that no patients who received chemo and progressed were radiated) 
- 100% combo price
- Discount rate: 3
- Data source for controls & exposed: Investigator assessment - Bouffet et al., NEJM 2023. 
- **Post-progression risk for subsequent progression = pre-progression risk for first progression in combo arm 
- 10,000 individuals
- 50 simulation runs

\newpage

### Investigator assessment, Targeted curve – Exponential, SoC curve – Exponential

```{r}
scen <- paste0("3_dr/","1a")
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase exp lifetime investigator combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Investigator assessment, Targeted curve – Log-Normal, SoC curve – Log-Normal

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase lnorm lifetime investigator combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Investigator assessment, Targeted curve – Log-Normal, SoC curve – Exponential

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase lnorm exp lifetime investigator combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Independent reviewer, Targeted curve – Exponential, SoC curve – Log-Normal

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase exp lifetime combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Independent reviewer, Targeted curve – Log-Normal, SoC curve – Log-Normal

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase lnorm lifetime combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

## Scenario #1c

- Combo vs. chemo as first-line treatment
- Lifetime duration of combo therapy
- 0% radiation at progression (Post-2014 SickKids data indicates that no patients who received chemo and progressed were radiated)
- **50% combo price reduction**
- Discount rate: 3
- Data source for controls & exposed: Investigator assessment - Bouffet et al., NEJM 2023.
- **Post-progression risk for subsequent progression = pre-progression risk for first progression in combo arm
- 10,000 individuals
- 50 simulation runs

\newpage

### Investigator assessment, Targeted curve – Exponential, SoC curve – Exponential

```{r}
scen <- paste0("3_dr/","1c")
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase exp lifetime investigator combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Investigator assessment, Targeted curve – Log-Normal, SoC curve – Log-Normal

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase lnorm lifetime investigator combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Investigator assessment, Targeted curve – Log-Normal, SoC curve – Exponential

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase lnorm exp lifetime investigator combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Independent reviewer, Targeted curve – Exponential, SoC curve – Log-Normal

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase exp lifetime combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Independent reviewer, Targeted curve – Log-Normal, SoC curve – Log-Normal

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase lnorm lifetime combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

## Scenario #1d

- Combo vs. chemo as first-line treatment
- Lifetime duration of combo therapy
- 0% radiation at progression (Post-2014 SickKids data indicates that no patients who received chemo and progressed were radiated)
- **75% combo price reduction**
- Discount rate: 3
- Data source for controls & exposed: Investigator assessment - Bouffet et al., NEJM 2023.
- **Post-progression risk for subsequent progression = pre-progression risk for first progression in combo arm
- 10,000 individuals
- 50 simulation runs

\newpage

### Investigator assessment, Targeted curve – Exponential, SoC curve – Exponential

```{r}
scen <- paste0("3_dr/","1d")
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase exp lifetime investigator combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Investigator assessment, Targeted curve – Log-Normal, SoC curve – Log-Normal

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase lnorm lifetime investigator combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Investigator assessment, Targeted curve – Log-Normal, SoC curve – Exponential

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase lnorm exp lifetime investigator combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Independent reviewer, Targeted curve – Exponential, SoC curve – Log-Normal

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase exp lifetime combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Independent reviewer, Targeted curve – Log-Normal, SoC curve – Log-Normal

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase lnorm lifetime combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

## Scenario #1e

- Combo vs. chemo as first-line treatment
- Lifetime duration of combo therapy
- 0% radiation at progression (Post-2014 SickKids data indicates that no patients who received chemo and progressed were radiated)
- 100% combo price
- Discount rate: 3
- **Data source for controls: n=34 SickKids controls**
- **Data source for exposed: Investigator assessment - Bouffet et al., NEJM 2023.**
- **Post-progression risk for subsequent progression = pre-progression risk for first progression in combo arm
- 10,000 individuals
- 50 simulation runs

\newpage

### Targeted curve – Exponential, SoC curve – Exponential

```{r}
scen <- paste0("3_dr/","1e")
fileloc <- here::here("04_analysis","output","scenarios",scen, "sickkids control invest exp lifetime combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Targeted curve – Log-Normal, SoC curve – Exponential

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "sickkids control invest lnorm lifetime combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

# 20% radiated at subsequent progression

## Scenario #1a

- Combo vs. chemo as first-line treatment
- Lifetime duration of combo therapy 
- 20% radiation at progression
- 100% combo price
- Discount rate: 1.5 
- Data source for controls & exposed: Investigator assessment - Bouffet et al., NEJM 2023. 
- **Post-progression risk for subsequent progression = pre-progression risk for first progression in combo arm 
- 10,000 individuals
- 50 simulation runs

\newpage

### Investigator assessment, Targeted curve – Exponential, SoC curve – Exponential

```{r}
scen <- paste0("rad/","1a")
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase exp lifetime investigator combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Investigator assessment, Targeted curve – Log-Normal, SoC curve – Log-Normal

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase lnorm lifetime investigator combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Investigator assessment, Targeted curve – Log-Normal, SoC curve – Exponential

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase lnorm exp lifetime investigator combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Independent reviewer, Targeted curve – Exponential, SoC curve – Log-Normal

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase exp lifetime combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Independent reviewer, Targeted curve – Log-Normal, SoC curve – Log-Normal

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase lnorm lifetime combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

## Scenario #1c

- Combo vs. chemo as first-line treatment
- Lifetime duration of combo therapy
- 20% radiation at progression
- **50% combo price reduction**
- Discount rate: 1.5
- Data source for controls & exposed: Investigator assessment - Bouffet et al., NEJM 2023.
- **Post-progression risk for subsequent progression = pre-progression risk for first progression in combo arm
- 10,000 individuals
- 50 simulation runs

\newpage

### Investigator assessment, Targeted curve – Exponential, SoC curve – Exponential

```{r}
scen <- paste0("rad/","1c")
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase exp lifetime investigator combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Investigator assessment, Targeted curve – Log-Normal, SoC curve – Log-Normal

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase lnorm lifetime investigator combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Investigator assessment, Targeted curve – Log-Normal, SoC curve – Exponential

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase lnorm exp lifetime investigator combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Independent reviewer, Targeted curve – Exponential, SoC curve – Log-Normal

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase exp lifetime combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Independent reviewer, Targeted curve – Log-Normal, SoC curve – Log-Normal

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase lnorm lifetime combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

## Scenario #1d

- Combo vs. chemo as first-line treatment
- Lifetime duration of combo therapy
- 20% radiation at progression 
- **75% combo price reduction**
- Discount rate: 1.5
- Data source for controls & exposed: Investigator assessment - Bouffet et al., NEJM 2023.
- **Post-progression risk for subsequent progression = pre-progression risk for first progression in combo arm
- 10,000 individuals
- 50 simulation runs

\newpage

### Investigator assessment, Targeted curve – Exponential, SoC curve – Exponential

```{r}
scen <- paste0("rad/","1d")
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase exp lifetime investigator combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Investigator assessment, Targeted curve – Log-Normal, SoC curve – Log-Normal

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase lnorm lifetime investigator combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Investigator assessment, Targeted curve – Log-Normal, SoC curve – Exponential

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase lnorm exp lifetime investigator combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Independent reviewer, Targeted curve – Exponential, SoC curve – Log-Normal

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase exp lifetime combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Independent reviewer, Targeted curve – Log-Normal, SoC curve – Log-Normal

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase lnorm lifetime combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

## Scenario #1e

- Combo vs. chemo as first-line treatment
- Lifetime duration of combo therapy
- 20% radiation at progression 
- 100% combo price
- Discount rate: 1.5
- **Data source for controls: n=34 SickKids controls**
- **Data source for exposed: Investigator assessment - Bouffet et al., NEJM 2023.**
- **Post-progression risk for subsequent progression = pre-progression risk for first progression in combo arm
- 10,000 individuals
- 50 simulation runs

\newpage

### Targeted curve – Exponential, SoC curve – Exponential

```{r}
scen <- paste0("rad/","1e")
fileloc <- here::here("04_analysis","output","scenarios",scen, "sickkids control invest exp lifetime combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Targeted curve – Log-Normal, SoC curve – Exponential

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "sickkids control invest lnorm lifetime combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

# 2-year treatment duration - Basecase

## Scenario #1a

- Combo vs. chemo as first-line treatment
- Lifetime duration of combo therapy 
- 0% radiation at progression (Post-2014 SickKids data indicates that no patients who received chemo and progressed were radiated) 
- 100% combo price
- Discount rate: 1.5 
- Data source for controls & exposed: Investigator assessment - Bouffet et al., NEJM 2023. 
- **Post-progression risk for subsequent progression = pre-progression risk for first progression in combo arm 
- 10,000 individuals
- 50 simulation runs

\newpage

### Investigator assessment, Targeted curve – Exponential, SoC curve – Exponential

```{r}
scen <- paste0("twoyr/","1a")
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase exp lifetime investigator combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Investigator assessment, Targeted curve – Log-Normal, SoC curve – Log-Normal

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase lnorm lifetime investigator combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Investigator assessment, Targeted curve – Log-Normal, SoC curve – Exponential

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase lnorm exp lifetime investigator combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Independent reviewer, Targeted curve – Exponential, SoC curve – Log-Normal

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase exp lifetime combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Independent reviewer, Targeted curve – Log-Normal, SoC curve – Log-Normal

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase lnorm lifetime combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

## Scenario #1c

- Combo vs. chemo as first-line treatment
- Lifetime duration of combo therapy
- 0% radiation at progression (Post-2014 SickKids data indicates that no patients who received chemo and progressed were radiated)
- **50% combo price reduction**
- Discount rate: 1.5
- Data source for controls & exposed: Investigator assessment - Bouffet et al., NEJM 2023.
- **Post-progression risk for subsequent progression = pre-progression risk for first progression in combo arm
- 10,000 individuals
- 50 simulation runs

\newpage

### Investigator assessment, Targeted curve – Exponential, SoC curve – Exponential

```{r}
scen <- paste0("twoyr/","1c")
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase exp lifetime investigator combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Investigator assessment, Targeted curve – Log-Normal, SoC curve – Log-Normal

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase lnorm lifetime investigator combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Investigator assessment, Targeted curve – Log-Normal, SoC curve – Exponential

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase lnorm exp lifetime investigator combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Independent reviewer, Targeted curve – Exponential, SoC curve – Log-Normal

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase exp lifetime combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Independent reviewer, Targeted curve – Log-Normal, SoC curve – Log-Normal

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase lnorm lifetime combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

## Scenario #1d

- Combo vs. chemo as first-line treatment
- Lifetime duration of combo therapy
- 0% radiation at progression (Post-2014 SickKids data indicates that no patients who received chemo and progressed were radiated)
- **75% combo price reduction**
- Discount rate: 1.5
- Data source for controls & exposed: Investigator assessment - Bouffet et al., NEJM 2023.
- **Post-progression risk for subsequent progression = pre-progression risk for first progression in combo arm
- 10,000 individuals
- 50 simulation runs

\newpage

### Investigator assessment, Targeted curve – Exponential, SoC curve – Exponential

```{r}
scen <- paste0("twoyr/","1d")
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase exp lifetime investigator combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Investigator assessment, Targeted curve – Log-Normal, SoC curve – Log-Normal

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase lnorm lifetime investigator combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Investigator assessment, Targeted curve – Log-Normal, SoC curve – Exponential

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase lnorm exp lifetime investigator combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Independent reviewer, Targeted curve – Exponential, SoC curve – Log-Normal

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase exp lifetime combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Independent reviewer, Targeted curve – Log-Normal, SoC curve – Log-Normal

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "basecase lnorm lifetime combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

## Scenario #1e

- Combo vs. chemo as first-line treatment
- Lifetime duration of combo therapy
- 0% radiation at progression (Post-2014 SickKids data indicates that no patients who received chemo and progressed were radiated)
- 100% combo price
- Discount rate: 1.5
- **Data source for controls: n=34 SickKids controls**
- **Data source for exposed: Investigator assessment - Bouffet et al., NEJM 2023.**
- **Post-progression risk for subsequent progression = pre-progression risk for first progression in combo arm
- 10,000 individuals
- 50 simulation runs

\newpage

### Targeted curve – Exponential, SoC curve – Exponential

```{r}
scen <- paste0("twoyr/","1e")
fileloc <- here::here("04_analysis","output","scenarios",scen, "sickkids control invest exp lifetime combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```

\newpage

### Targeted curve – Log-Normal, SoC curve – Exponential

```{r}
fileloc <- here::here("04_analysis","output","scenarios",scen, "sickkids control invest lnorm lifetime combo")
res_model <- list.files(paste0(fileloc, "/res_model"),full.names = TRUE)[index]
```

```{R}
create_table(res_model)
```

```{r, out.width="100%"}
plots <- create_trace(fileloc)
plots$trace_plot
```

```{r, out.width="100%"}
grid.arrange(plots$ae_plot1, plots$ae_plot2, nrow=2)
```

```{r, out.width="100%"}
grid.arrange(plots$surv_plot1, plots$surv_plot2, plots$surv_plot3, nrow=3)
```